---
title: "Analysis of bicycle rentals\nin Washington D.C. and Seoul"
author: "Student 200019804"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
library(magrittr)
library(lubridate)
library(patchwork)
library(knitr)
library(GGally)
setwd("~/R/R WD")
```

```{r Reading in data, echo=FALSE}
setwd("~/R/R WD")
seoul <- read.csv("BikeSeoul.csv")
washington <- read.csv("BikeWashingtonDC.csv")
```

```{r data wrangling, echo=FALSE}

seoul2 <- seoul %>%
  
  select(-Visibility..10m.,
         -Dew.point.temperature.C.,
         -Solar.Radiation..MJ.m2.,
         -Rainfall.mm., -Snowfall..cm.) %>%
  
  filter(Functioning.Day == "Yes") %>%
  
  select(-Functioning.Day)


colnames(seoul2) <- c("Date", "Count", "Hour", "Temperature", "Humidity",
  "WindSpeed", "Season", "Holiday") #way to do with pipe????


seoul3 <- mutate(seoul2, "Date" = as.Date(Date, format = "%d/%m/%Y")) %>%
  
  mutate("FullDate" = make_datetime(
           
           year(Date),month(Date), day(Date),Hour))



seoul3$Holiday <- as.factor(seoul3$Holiday)
levels(seoul3$Holiday) <- c("Yes", "No")


seoul3$Season <- as.factor(seoul3$Season)
seoul3$Season <- factor(seoul3$Season, 
                        levels = c("Spring","Summer",
                                   "Autumn", "Winter"))


washington2 <- washington %>%
  
  select(-instant, -yr, -mnth, -weekday, -workingday,
         -weathersit, -atemp, -casual, -registered)

colnames(washington2) <- c("Date", "Season", "Hour", "Holiday",
                "Temperature", "Humidity", "WindSpeed", "Count")

washington3 <- washington2 %>%
  
  mutate("Humidity" = Humidity*100,
         "Temperature" = 47*Temperature - 8,
         "WindSpeed" = (5/18)*67*WindSpeed,
         "Date" = as.Date(Date, format = "%Y-%m-%d"),
         "FullDate" = make_datetime(
           year(Date),month(Date), day(Date),Hour)) 
  

washington3$Season <- factor(washington3$Season,
                             levels = c("2", "3","4", "1")) # re-ordering

levels(washington3$Season) <- c("Spring", "Summer", # re-naming
                                "Autumn", "Winter")
  

washington3$Holiday <- factor(washington3$Holiday, #re ordering
       levels = c("1", "0"))

levels(washington3$Holiday) <- c("Yes", "No") # re naming

```

## Introduction

In this project we investigate bicycle rentals in two cities, Seoul, South Korea, and Washington D.C., USA by examining locally collected data sets and through the use of statistical modelling. 

\newpage

## Climate

To begin to understand bike rental numbers in the two cities, it is important to understand each city's climate.

```{r, temperature plots, echo=FALSE}
kr1 <- ggplot(seoul3, aes(x = Date, y = Temperature)) +
  geom_point(aes(color = Hour), alpha = 0.2) +
  
  geom_smooth(color = "black") + theme_light() +
  
  labs(x= NULL, y = "Air Temperature (°C)", title = 
         "Air temperature from December to November",
       subtitle = "Seoul 2017/18") +
  
  scale_color_gradient2(mid = "#66C9FF",low = "#001B6B",
                        high = "#001B6B", space = "Lab",
                        midpoint = 12) + 
  
  expand_limits(y = c(-20, 40)) + 
  
  scale_y_continuous(breaks = seq(-20, 40, 10)) 


dc1 <- ggplot(washington3[7905:16637,], aes(x = Date, y = Temperature)) +
  
  geom_point(aes(color = Hour), alpha = 0.3) + 
  
  geom_smooth(color = "black") + theme_light() +
  
  labs(x= NULL, y = "Air Temperature (°C)",
   subtitle = "Washington D.C. 2011/12") +
  
  scale_color_gradient2(mid = "#FFA600",low = "#6B0000",
                        high = "#6B0000", space = "Lab",
                        midpoint = 12) + 
  
  expand_limits(y = c(-20, 40)) + 
  
  scale_y_continuous(breaks = seq(-20, 40, 10))


```

```{r fig1, fig.width=5,fig.height=5,fig.cap="\\label{fig:fig1} Air Temperature in Seoul and Washington D.C.", message=FALSE, fig.align='center', echo=FALSE}
kr1/dc1
```

Figure \ref{fig:fig1} shows the air temperature (°C) throughout a year in Washington D.C. and Seoul. The temperature peaks between July and August in both cities with Seoul experiancing a shorter, more pronounced period of peak temperature. Visibly the climate in Seoul is more extreme than in Washington D.C. in the chosen year, with a hotter summer and colder winter.

```{r, echo=FALSE}
temps <- data.frame("Max" = c(max(seoul3$Temperature),
                                max(washington3$Temperature)),
                   "Min" = c(min(seoul3$Temperature),
                                 min(washington3$Temperature)),
                   "Mean" = c(mean(seoul3$Temperature),
                                 mean(washington3$Temperature)))

rownames(temps) <- c("Seoul", "Washington D.C.")

kable(temps, caption = "Temperature (°C)")
```

Seoul experiences a slightly higher maximum temperature and a much colder minimum temperature, resulting in Washington D.C. having the higher mean temperature.

\newpage

### Bicycle Rentals

```{r, echo=FALSE, message=FALSE, include=FALSE, eval=FALSE}

count_month_DC <- washington3 %>% group_by(Season) %>%
  
  summarise("Count" = mean(Count))

count_month_KR <- seoul3 %>% group_by(Season) %>%
  
  summarise("Count" = mean(Count))


dc2 <- ggplot(count_month_DC, aes(x = Season, y = Count, fill = Season)) +
  
  geom_bar(stat = "identity", color = "black") + theme_light() +
  
  theme(legend.position='none') + expand_limits(y = 250) + 
  
  scale_y_continuous(breaks = seq(0, 250, 50)) +
  
  labs(x = NULL, y = NULL,
       subtitle = "Washington D.C.")



kr2 <- ggplot(count_month_KR, aes(x = Season, y = Count, fill = Season)) +
  
  geom_bar(stat = "identity", color = "black") + theme_light() +
  
  theme(legend.position='none') + expand_limits(y = 1100) + 
  
  scale_y_continuous(breaks = seq(0, 1100, 200)) +
  
  theme(plot.title = element_text(size=15)) +
  
  labs(x = NULL, y = NULL, subtitle = "Seoul", title = 
         "Average number of bike rentals in each season",)

```

```{r, echo=FALSE}
wash_box <- ggplot(washington3, aes( x = Season, y = Count)) +
  
  geom_boxplot(aes(fill = Season)) + theme_light() +
  
  labs(subtitle = "Washington D.C.", x = NULL, y = "Rentals per day") +
  
  theme(legend.position='none')

seoul_box <- ggplot(seoul3, aes( x = Season, y = Count)) +
  
  geom_boxplot(aes(fill = Season)) + theme_light() +
  
  labs(title = "Bicycle rentals per day by season", subtitle = "Seoul", x = NULL,
       y = "Rentals per day") +
  
  theme(legend.position='none')
```

Figure \ref{fig:fig2} shows the impact of Seoul's colder winter on bike rentals. The bike rentals in winter are noticeably lower than those in Washington D.C. when compared to the other seasons.

Figure \ref{fig:fig2} also shows that the bike rentals in each season are related to the temperature with the highest average rentals in summer, the hottest month for both cities, and the lowest in winter. The bike rentals vary less in Washington D.C. than in Seoul, this is likely due to Washington's less extreme climate exhibited in Figure \ref{fig:fig1}.


```{r fig2, fig.width=4,fig.height=4,fig.cap="\\label{fig:fig2} Bike rentals in Seoul and Washington D.C.", message=FALSE, fig.align='center', echo=FALSE}

seoul_box / wash_box
```


\newpage

```{r, echo=FALSE, message=FALSE}
count_time_DC <- washington3 %>% group_by(Season, Hour) %>%
  
  summarise("Count" = mean(Count))


count_time_KR <- seoul3 %>% group_by(Season, Hour) %>%
  
  summarise("Count" = mean(Count))




DC_time <- ggplot(count_time_DC, aes(x = Hour, y = Count)) + 
  
  expand_limits(x = 22) + 
  
  scale_x_continuous(breaks = seq(0, 22, 2)) +
  
  geom_smooth(aes(color = Season), se = F) + theme_light() +
  
  labs(subtitle = "Washington D.C.", y = "Average Rentals")



KR_time <- ggplot(count_time_KR, aes(x = Hour, y = Count)) + 
  
  expand_limits(x = 22) + 
  
  scale_x_continuous(breaks = seq(0, 22, 2)) +
  
  geom_smooth(aes(color = Season), se = F) + theme_light() +
  
  labs(title = "Average bike rentals per hour", subtitle = "Seoul",
       x = NULL, y = "Average Rentals") +
  
  theme(legend.position='none')
```

```{r fig3, fig.width=4,fig.height=4, message=FALSE, fig.align='center', echo=FALSE, fig.cap="\\label{fig:fig2} Placeholder"}

KR_time / DC_time

```

Figure \ref{fig:fig3} shows the relationship between time of day and bike rentals in each season in the two cities. In Washington D.C. the average bike rentals peak between 16:00 and 17:00 year round, with summer peaking the latest.

Unlike in Washington D.C., the average bike rentals per hour in Seoul are very different for each season. Firstly the the average bike rentals in Seoul peak later in the day than in Washington D.c., most noticeably in summer and autumn peaking at 20:00 and 18:00 respectively. Visibly the the average bike rentals in Seoul stay at their maximum values for longer than in D.C..



```{r,  eval=FALSE,echo=FALSE, message=FALSE}

seoul_DC <- rbind(mutate(washington3, 
                         City = rep("Washington D.C.", nrow(washington3))),
                  mutate(seoul3, 
                         City = rep("Seoul", nrow(seoul3)))) %>% 
  
  group_by(Season, Hour, City) %>%
  
  summarise("Temp" = mean(Temperature))


ggplot(seoul_DC, aes(y = Temp, x = Hour)) + geom_smooth(aes(color = City, type = Season))

```

```{r, echo=FALSE, message=FALSE}
count_holiday_DC <- washington3 %>% group_by(Holiday) %>%
  
  summarise("Count" = mean(Count))

count_holiday_KR <- seoul3 %>% group_by(Holiday) %>%
  
  summarise("Count" = mean(Count))
```


```{r, echo=FALSE}
kable(count_holiday_KR, caption = "Holiday Bike Rentals in Seoul")

```

Table X and Y show the impact of holidays on bike rentals in Seoul and Washington D.c. respectively. On average, the bike rentals in Seoul on a holiday are 28% lower than a non-holiday. This is likely due to the lack of commuters renting bikes.

Similarly, the average bike rentals in Washington D.C. on a holiday are 17% lower than a non-holiday. 

The difference between bike rentals on holidays in the two cities suggests a cultural difference in the way bicycles are used for leisure and work.

\newpage

```{r, echo=FALSE}
kable(count_holiday_DC, caption = "Holiday Bike Rentals in Washington D.C.")
```

### Meteorological effects

Next we can analyse the effect of different meteorological factors on the number of bike rentals in the two cities. 

```{r, echo=FALSE}
kr_met1 <- ggplot(seoul3, aes(y = Count, x = WindSpeed)) +
  
  geom_jitter(width = 0.02, height = 0,
              alpha = 0.5,color = "#001B6B") + 
  
  geom_smooth() +
  
  labs(x = "Wind Speed (m/s)", title =
  "Number of bike rentals for different meteoroloigcal\nfactors in Seoul")

kr_met2 <- ggplot(seoul3, aes(y = Count, x = Temperature)) +
  
  geom_point(alpha = 0.5, color = "#001B6B") + 
  
  geom_smooth() + labs(x = "Temperature (°C)")



kr_met3 <- ggplot(seoul3, aes(y = Count, x = Humidity)) +
  
  geom_point(alpha = 0.5, color = "#001B6B") + 
  
  geom_smooth() + labs(x = "Humidity (%)")

#this one sucks


dc_met1 <- ggplot(washington3, aes(y = Count, x = WindSpeed)) +
  
  geom_jitter(width = 0.1, height = 0,
              alpha = 0.5,color = "#001B6B") +
  
  geom_smooth() +
  
  labs(x = "Wind Speed (m/s)", title =
         "Number of bike rentals for different meteoroloigcal\nfactors in Washington D.C.")

dc_met2 <- ggplot(washington3, aes(y = Count, x = Temperature)) +
  
  geom_point(alpha = 0.5, color = "#001B6B") + 
  
  geom_smooth() + labs(x = "Temperature (°C)")



dc_met3 <- ggplot(washington3, aes(y = Count, x = Humidity)) +
  
  geom_point(alpha = 0.5, color = "#001B6B") + 
  
  geom_smooth() + labs(x = "Humidity (%)")
```

```{r fig5, fig.width=5,fig.height=3,fig.cap="\\label{fig:fig5} Seoul bike rentals", message=FALSE, fig.align='center', echo=FALSE}

kr_met1 / (kr_met2 + kr_met3)
```

Figure \ref{fig:fig5} shows the impact of air temperature, humidity, and wind speed on bicycle rentals in Seoul. Bike rentals increase with air temperature up to 30°C before starting to decrease, suggesting people in Seoul are more likely to rent a bicycle on warm days but not on the hottest days. This is expected as cycling in low temperatures presents dangers like ice as well as exposure to the cold temperature itself. 

The impact of wind speed appears to be lesser, only showing a small decrease for speeds above 4m/s, showing that only high wind speeds have any meaningful effect on the number of bike rentals.

Humidity?

```{r fig6, fig.width=5,fig.height=3,fig.cap="\\label{fig:fig6} Washington D.C. bike rentals", message=FALSE, fig.align='center', echo=FALSE}

dc_met1 / (dc_met2 + dc_met3)
```

The impact of the three meteorological factors is similar in Washington D.C., shown in Figure \ref{fig:fig6}. One noticeable difference is that the bike rentals in Washington D.C. do not see such a decline at the higher temperatures close to 40°C. The decline in bike rentals in Washington D.C. with increasing humidity is more pronounced than in Seoul, starting at ~25% compared to ~50% in Seoul. 
\newpage
\newpage

## Modelling

In order to model and predict bike rentals in Seoul and Washington D.C. we apply linear regression to the logarithm of the number of bicycle rentals, $log(Count)$, with season, air temperature, humidity, and wind speed as independent variables. The model can be represented as

$$
log(Count)\sim Season\;+\; Temperature \;+\; Humidity \;+\; Wind Speed.
$$

```{r, echo=FALSE}
fit_kr <- lm(log(Count) ~ Season + Temperature +
               
               Humidity + WindSpeed, data = seoul3)



fit_dc <- lm(log(Count) ~ Season + Temperature +
               
               Humidity + WindSpeed, data = washington3)


```

```{r, echo = FALSE}
kr_resid <- resid(fit_kr)
dc_resid <- resid(fit_dc)
```

### Checking Model Assumptions

To assess the effectiveness, reliability, and validity of the models, it is important to check the assumptions about both sets of data. Firstly we check the assumption of normality of residuals for both models.

Figures \ref{fig:fig7} and \ref{fig:fig8} show Normal Quantile-Quantile plots for the residuals of the models for Seoul and Washington D.C. respectively.

```{r fig7, fig.width=4,fig.height=3,fig.cap="\\label{fig:fig7} Seoul normal Q-Q plot", message=FALSE, fig.align='center', echo=FALSE}
qqnorm(kr_resid, main = "Seoul Normal Q-Q Plot")

abline(a = 0, b = 1, col = "blue", lty = 2)
```

The data for Seoul, shown above in Figure \ref{fig:fig6}, exhibits a slight left-skew for its tails however, the main body of the data lies on the reference line, indicating the assumption of normality of residuals is fair.

```{r fig8, fig.width=4,fig.height=3,fig.cap="\\label{fig:fig8} Washington D.C. normal Q-Q plot", message=FALSE, fig.align='center', echo=FALSE}
qqnorm(dc_resid, main = "Washington D.C. Normal Q-Q Plot")

abline(a = 0, b = 1, col = "red", lty = 2)
```

The data for Washington D.C.,Figure \ref{fig:fig7}, is light tailed although the upper part of the data lies well on the reference line, again suggesting normality of residuals is a fair assumption.

\newpage

```{r fig9, fig.width=5,fig.height=3,fig.cap="\\label{fig:fig9} Seoul scatterplot matrix", message=FALSE, fig.align='center', echo=FALSE}

ggpairs(select(seoul3,Temperature, Humidity, WindSpeed))

```

Figure X shows that the three meteorological factors selected as predictors exhibit strong multicollinearity, meaning the assumption of independence between variables isn't shown in the data, which may negatively effect the accuracy and validity of the model for Seoul.

The data for Washington D.C., shown in Figure Z, suffers from the same issue.

```{r fig10, fig.width=4,fig.height=3,fig.cap="\\label{fig:fig10} Washington D.C. scatterplot matrix", message=FALSE, fig.align='center', echo=FALSE}

ggpairs(select(washington3,Temperature, Humidity, WindSpeed))

```


\newpage

### Model Analysis

```{r, echo=FALSE}
summary(fit_kr)
```

For the Seoul model, all independent variables are statistically significant at the 1% level, other than than summer not being significantly different to the intercept, spring. A multiple R-squared value of 0.49 indicates that the model is not a very good fit as there is still much variation in bike rentals not explained by the predictors. This may be due to the string multicollinearity between the predictors, and important factors such as time of day and holiday status being left out of the model.

```{r, echo=FALSE}
summary(fit_dc)
```

For the Washington D.C. model, all predictors are statistically significant at the 1% level. The low multiple R-Squared value of 0.28 suggests the model is not a good fit for the data, for similar reasons as the Seoul model.

### Coefficients 

```{r kr_tab, echo=FALSE}

kr_conf <- confint(fit_kr, level = 0.97)

rownames(kr_conf) <- c("Intercept", "Summer", "Autumn",
                       "Winter","Temperature", "Humidity", "Wind Speed")

kable(kr_conf, caption = "Seoul model parameters 97% CI")

```

```{r dc_tab, echo=FALSE}
dc_conf <- confint(fit_dc, level = 0.97)

rownames(dc_conf) <- c("Intercept", "Summer", "Autumn",
                       "Winter","Temperature", "Humidity", "Wind Speed")

kable(dc_conf, caption = "Washington D.C. model parameters 97% CI")
```

## Prediction

```{r, echo=FALSE}
params <-data.frame(Temperature = 0,
                  WindSpeed = 0.5,
                  Humidity = 20,
                  Season = "Winter")

```

```{r, echo=FALSE}

pred_kr <- predict(fit_kr, newdata = params,
        
        interval = "confidence", level = 0.9)

colnames(pred_kr) <- c("Prediction", "5%", "95%")

pred_kr <- round(exp(pred_kr), 1)

pred_dc <- predict(fit_dc, newdata = params, 
        
        interval = "confidence", level = 0.9)

colnames(pred_dc) <- c("Prediction", "5%", "95%")

pred_dc <- round(exp(pred_dc), 1)

preds <- rbind(pred_kr, pred_dc)

rownames(preds) <- c("Seoul", "Washington D.C.")

```

Using the models for Seoul and Washington, we can make predictions for the average number of bicycle rentals per hour on a unspecified day in a given season with given temperature, humidity, and wind speed. We predict for freezing (0°C) temperature, with 0.5m/s wind speed, humidity of 20%, and in winter.                   

```{r,echo=FALSE}

kable(preds, caption = "Bike rental prediction")

```

The model predicts 370 bike rentals per hour in Seoul and 72 in Washington D.C..
